\chapter{Introduction}
\label{chap:intro}
  Mobile robotics is nowadays a~prestigious research area, and building a~highly functional 
  mobile robot is considered to be difficult. 
  Not only is it considered difficult, but it is a~fascinating process, which
  involves a~lot of engineering disciplines.
  
  %The processing power of microchips has grown more than ten times in~the~last decade. 
  %A large variety of different sensors are available on~the~market.	
  A lot of cheap and sufficient components result in~a~boom of simple robots used in~daily life. 
  Lawnmowers and vacuum cleaners are typical examples of mobile robots invasion~into our households in~these days.
  We also get used to robot prototypes, which are highly specialised in~the~space exploration~or in~army services.

  Despite the~rapid hardware development and robots increasing popularity,
  the~research in~the~field of mobile robots is at its beginning.
  The robots dominates in industry and they are already useful in our daily
  life, but the true real life problems are still difficult for them.
  Many complex problems like driving a car in full traffic or to move in humam
  like environment are still a real challenge for a robot.
   
  The young scientests often are discouraged by
  studying hardware details during a~building of their own robot.
  Luckily, several educational robots were made to help the novices.
  Educational robots help students focus on the other aspects of mobile robotics than the hardware one.
  
  This work describes and implements a~library for one of the~educational robots.
  Its name is e-Puck and it is a~typical example of an educational robot for students at the~university level. 
  E-Puck has a~clean, simple and robust mechanical design, which is easy to understand.
  Bluetooth wireless communication~enables sending data~between e-Puck and PC.
  A camera, eight infra~red (IR) sensors,	an accelerometer, encoders and three microphones 
  are enough for a~robot to feel the~real world.
  E-puck robot can reply to any type of perception~by performing several actions. 
  The robot is able to emit light from light emitting diodes (LEDs) or to play a~sound.
  Two differential stepper motors facilitate a~precise movement. 
  
  Eight of e-Puck LEDs are located on~its perimeter. There are red. 
  Four of green LEDs are placed in~its translucent body and
  one stronger front LED is located next to the~camera. 
  The front LED illuminates the~terrain~in~front of the~robot to
  take better pictures. The rest of LEDs together with the~speaker are usually used to
  generate feedback. 
  
  E-puck's sensors and actuators allows to solve a~large scale of problems from the~field of mobile robotics. 
  On the~other hand, the~processing power of e-Puck does not allow to 
  process all outputs from its sensors. For example, resolution~from e-Puck's camera~is 640 * 480 pixels,
  but the~pictures taken from the~camera~are trimmed under 50 * 50 pixels, because there is a~lack of space in~
   e-Puck's microchip memory.
   
   E-puck robot and its sensors and actuators can be controlled only by a~low level C code,
    which needs to be downloaded to e-Puck's memory, and has to be run by the~robot's microchip.
   On contrary, a~program downloaded to e-Puck can communicate with
   a~PC and it can let the~PC to control the~e-Puck robot. 
   The program has to define a~protocol, in~which it communicates between e-Puck and PC over Bluetooth.
  On e-Puck such a~program exists and is called {\it BTCom}. 
  {\it BTCom} is distributed under an open source licence with e-Puck.

  Programming e-Puck's microchip needs knowledge of sensors and actuators interfaces.
  Furthermore debugging is limited to blinking with diodes or playing sounds.
  Mentioned reasons and insufficient processing power hinder rapid development of even a~simple program, 
  which makes e-Puck programming inconvenient for students.
  
  At the~time of writing this thesis there was only one well developed software for e-Puck called Webots,
  which makes programming e-Puck easier.
  Webots is an useful develop and simulation~environment, but it is an expensive commercial software.
  Other simulators, which can be free alternatives for Webots, are introduced in~Chapter ~\ref{chap:software}.
  
  %todo uvest proc rikam o remote control
  The drawbacks of low level programming are removed by the~{\it Elib} library, which is the~main~
  contribution~of this work.
  {\it Elib} library uses {\it BTCom} program on~e-Puck.
  {\it Elib} library runs on~PC and communicates with the~robot via~Bluetooth. 
  A program, which uses {\it Elib} library,	makes an e-Puck a~carrier of sensors. It only gives e-Puck 	
  commands to change state of its actuators, whereas the~whole algorithmic code runs on~PC.
  
  On the~other hand, the~implementation~of {\it Elib} brings new problems.
  This thesis introduces the~{\it Elib's} problems and discusses 
  under which conditions is the~approach of {\it Elib} suitable.	
  \subsection*{Thesis structure}
  We introduce a~position~of mobile robotics in~the~contemporary world in~Chapter ~\ref{chap:robotics}.

  Chapter ~\ref{chap:epuck} describes e-Puck design and presents its drawbacks and benefits.

  Chapter ~\ref{chap:software} lists software which help programmers to develop applications for robots.
  In this chapter the~usage of a~software for e-Puck is discussed. 
  
  Chapters ~\ref{chap:elib} and ~\ref{chap:usage} presents the~design and usage of {\it Elib} library in~detail.
  Conclusion~sums up the~properties of {\it Elib}.
  A~simple installation~guide of {\it Elib}, a list of {\it Elib's} functions and a list of {\it Elib's} exceptions  are located in enclosed Appendixes.
  An installation and a~user guide for {\it Elib Joystick} application~is also included. 


