r{Introduction to mobile robotics}
\label{chap:robotics}
	Human beings always created tools to make its job easier. 
	Even the word "robot", first used by Karel Capek in a novel RUR published in 1920, came from world "robota", 
	which in Czech means labour.
	\input{spirit_rover.TpX}%Rover Spirit on Mars
	Nowadays mobile robots are useful in many areas. Robots excel in exploration of danger places. 
	For example rover Spirit and Delta II flew to Mars.
	Delta II is a space rocket. On the other hand Spirit, is a robot of our imagination. It is a six wheel robot
	which is similar to police robots for a bomb manipulation.
	Robots like Delta II and Spirit allow humankind to explore so distant places, which are too far for a man to travel there.
	
	Mobile robots are not narrow specialised in adventurous actions and impossible missions. 
	Their main contribution is in daily life.
	In many cities of the world the robots are used in public transport. They drive the trains in metro,
	help pilots to take off or land planes. They are used to park cars or they control cruise controls.
	Solutions which were studied in mobile robotics are usually incorporated in daily life use. 
	One time a car computer helps to park a car,
	another time a vacuum cleaner tidies up a room.
	Robots are useful not only as independent machines. Most of the robots help people with difficult task, 
	but people control their actions, which allow to people keep robots under control.
	Nice example is a parking robot present in many contemporary cars.
	
	
	On the other hand, the research in mobile robotics tries to produce robots which are autonomous as much as possible.
	The developers take all the responsibility for a robot, which is particularly demanding if the robot interacts with people. 
	There are areas where the research and industry already succeed.
	
	Autonomous lawnmowers and vacuum cleaners are broadly used in households and in garden and no one has fear from them.
	A number of robotic toys have been already introduced.
	A robotic dog Aibo introduced in 1999 is a nice example of a toy robot.
	\input{aibo.TpX}
	\begin{quote}
	"About 4.4 million units for domestic use and about 2.8 million units for entertainment and leisure sold up to end 2009."
	\cite{worldrob}
	\end{quote}
	
	Not only entertainment robots are designed for human-robot interaction. In Japan there is a big stress 
	on creating	robots for health care, which will compensate lack of medical personnel. 
	A robot has to be a humanoid, because especially children and older people are used to 
	communicate with other people and not with machines.
	
	Professor Hiroshi Ishiguro reached a great success in field of humanoid robots. 
	He created several generations of humanoid robots.
        Among other humanoids there are copies of him and his daughter.
	The last generation of robots are so similar to humans, that they are almost unrecognisable.
	Hiroshi Ishiguro is concentrating on simulating human behaviour, speech and motion, 
	but he also invented a touch sensor, which substitus human skin on his robots.
	\input{ishiguro_clone.TpX}%Ishiguro with his geminoid
	
	The humanoid robots and also the Aibo robot are not  wheeled robots, although the wheeled 
	robot systems are still	dominating. The reason is in the field of the activity of the robot. 
	The entertainment and the health care robots are in the human environment. 
	Robots avoid a lot of different obstacles and they have to use stairs.
	The industrial robots are usually indoors, where the terrain is adapted to machines movement. 
	Also the agricultural robots have convenient conditions for wheel movement.
	
	Not only the design of robots, but also the control of robots differs and evolves.
	
	In the eighties the sense-plan-act model was current.
	Later in the nighties the behaviour model was introduced by Brooks,
	which was a very modern and sometimes strictly implemented.	
	After a decade a compromise between the behaviour based robotics and the simple planning was popular.
	In the nighties neural networks and other modern, biologically motivated attitudes were first introduced.
	Genetic algorithms and neural networks have enormous success in solving complicated control problems like motion of
	robot with a number of legs, implementing very sophisticated behaviour and so on.
	
	To conclude, mobile robotics is a dynamically evolving science,
	which has all preconditions to be one of the leading research areas in the twenty first century.
	The value of the market increased to 6.2 billion dollars in 2009 according to the International Federation of Robotics.
	\cite{worldrob}
	\begin{quote}	 	
	"I can envision a future in which robotic devices will become a nearly ubiquitous part of our day-to-day lives," 
	\end{quote}
	says Bill Gates, who was at the beginning of the computer revolution. %todo citation?
	

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{E-Puck}
\label{chap:epuck}
	An object of interest of a this thesis is a programming of the e-Puck robot and e-Puck itself.
	In this chapter e-Puck robot and its design is introduced. We focus on a programmer's point 
	of view of e-Puck qualities.
	For each sensor its functions, acceptable values and typical use is shortly presented. 
	After mentioning the e-Puck history the chapter continues with describing e-Puck's mechanical design and stepper motors. 
	The chapter finishes with a description of a e-Puck's camera and outlining e-Puck possibilities.
\section{Origin of e-Puck}
	E-Puck was developed in summer of 2004 at the École Polytechnique Fédérale de Lausanne (EPFL) 
	as an open tool. E-Puck designers used open software and hardware development model. 
	Motivation for a new robot arose from an absence of a very small educational robot,
	which is sufficiently efficient, and can be used for education in many research areas.
	E-Puck can be applied in automatic control, signal processing or 
	distributed intelligent system research. E-Puck's structure is robust and simple to care about, 
	because e-Puck is intended to be used by students.
	Designers of e-Puck tried to use manufacturing components as much as possible 
	in order to keep the price low.		 
	
	The first generation of e-Pucks was replaced by the second stable generation from 2006.
	So far, more than 2000 robots have been manufactured.
	There are several distributors for Switzerland and America, one is in Japan
	\footnote{\url{http://www.roadnarrows.com/robotics/}, \url{http://www.aai.jp/}, \url{http://www.k-team.com/},
	\url{http://www.cyberbotics.com/products/robots/e-puck.html}, \url{http://www.gctronic.com/},
	\url{http://www.aai.ca/robots/e-puck.html}}.
	The price of e-Puck should be between 450 and 550 euro.%todo znak eura

\section{Mechanical Design}
	\input{epuck_str.TpX} %pouzite \cite{robotica2009}
	The mechanical design  is shown in Figure ~\ref{pic:epuck_str}.
	E-Puck consists of a rounded transparent body, a battery, two stepper motors with
	wheels on their axis, a printed circuit board, plastic ring of LEDs, 
	a camera and a default extension board.
	There are extensions like floor sensors, a rotating scanner or a turret with linear cameras,
	which can scale up e-Puck's sensors. If we are talking about e-Puck in this thesis,
	it is meant e-Puck only with the default extension.
	
	A default robot is more than 60 mm 
	high and it has a diameter of 75 mm. 
	The battery is placed in the bottom and can be easily extracted and recharged separately in a charger.
	Two stepper motors with wheels are screwed to a plastic body and are located on axis of e-Puck
	in order to allow robot turn around on place. Wheels has a diameter of 41 mm and the perch is 53 mm.
	Printed circuit board is fixed to the top of the body and there is ring of LEDs' around the board.
	A camera is placed on the front side of robot lying on the axis between the wheels.
	The extension board covers the main printed circuit board.
\section{Sensors, actuators and heart of a robot}
	Sensors and actuators determine the possibility of the robot usage.
	Luckily, e-Puck has a lot of sensors of different kinds and is equipped with typical actuators.
	The following paragraphs shortly describes e-Puck's sensors and actuators and mention a few problems 
	with transmitting data between the devices and PC using {\it BTCom}. {\it BTCom} protocol is used by {\it Elib},
	so {\it Elib} is to some extent dependent on {\it BTCom}.
	The actuators consist of motors, 8 red LEDs on perimeter, 4 green body LEDs,
	 which are turned off and on together, a front LED and a speaker.
	 
	Stepper motors are a great advantage of e-Puck, because the motion of wheel can be split
	into small steps. One wheel revolution corresponds to 1000 steps of motor. The 
	diameter of the wheel is 41 mm. If the wheel makes revolution, the wheel goes 128.8 mm. 
	In conclusion a thousand of steps matches 128.8 mm and one step is 0.128 mm. 
	Motors are equipped with encoders and the motors are very accurate, 
	which in combination with simple odometry really helps in localisation tasks.
	A nice feature of encoders is that their value can be set at any time.
	The maximum speed of the motors is one revolution per second in both directions.
	{\it BTCom} allows a programmer to set the speed of motors and also get and set the values
	of encoders.
	 
	All e-Puck LEDs can be used for debugging or for making robot visible to other devices.
	Furthermore, the front LED can be used to illuminate the terrain in front of  e-Puck.
	Each LED can be turned off, turned on or set to an inverse state.
	
	Three axis accelerometers are placed inside the robots body. In the rest position
	the accelerometers measure the slant of e-Puck. They can also measure acceleration of e-Puck
	and for example detect a collision or a falling state of e-Puck. 
%	There is no problem with using accelerometer via {\it BTCom}.
	
	Infra Red (IR) sensors are typical sensors for mobile robotics. E-Puck has eight of them.
	Four are located in the front part of e-Puck, two are on both sides, two on back side.
	Sensors work in two modes. First they measure ambient infra red light.
	In the second mode IR sensors emit IR light and they measure reflected light, so 
	they can determine an obstacle, if it is near enough.
	Both functions are available in {\it BTCom}. 				
	Calibration of sensors increases the precision of detecting near obstacles.
	Infra red sensors are capable of recognising an obstacle under 4 cm.
	
	The Speaker together with microphones is a suitable communication tool between the people and e-Puck.
	On the other hand, a limited processing power of e-Puck does not allow to store and
	 play complicated sounds. It is also impossible to use a speaker with microphones to speech
	 recognition due to insufficient processing power.
	Despite the limitation of processing power the microphones can be easily used to locate
	the source of sound via amplitude measurement, because the microphones are placed near the perimeter
	in a triangle. Microphones measure current amplitude of sound. As exact distances between microphones 
	are known, we can compute frequency of the sound using Fast Fourier Transformation (FFT).
	Digital Signal Processor (DSP) is suitable for computing FFT,
	 which will be introduced in this chapter.
	Maximal acquisition speed is 33 kHz. 
	For more information see \cite{sound}.
	
	A program which uses {\it BTCom} can capture values of amplitude only in irregular intervals, so 
	it is not possible to compute frequency of sound by running FFT on Personal Computer. 
	However to locate the source of sound is still possible.
	
	\input{epuck.TpX}
	%todo what is zoom? does it only narrow field of view? or does it make the picture look closer?
	E-Puck's camera is a colour CMOS camera with 640 * 480 resolution. 
	Because only 8 kB of RAM memory is available and a part of it is used by running program,
	the picture size has to be reduced in order to save the image in a memory.
	The image processing is really demanding on the processing power so it is not 
	convenient to be run on the slow e-Puck's processor.
	
	On the other hand, using {\it BTCom} solves the problem with limited processing power
	by sending picture to Personal Computer (PC). PC has enough resources to process the image fast,
	 but the transport of an image takes a long time too. 
	 For example capturing and sending a colour image of size 40 * 40 pixels 
	takes more than 0.2 seconds, if it is sent over Bluetooth.
	{\it BTCom} can set height, width, colour mode and zoom. A colourful picture is twice as big as the same picture taken
	in grey scale mode.
	 Zoom has three acceptable values 1, 4 and 8. One is for the biggest zoom, 8 represents the smallest.
	Width and height are limited only by the size of the available memory in e-Puck.
	
	Processor, dsPIC 30F6014, is the heart of e-Puck and runs at 60 MHz, which correspond to 15 MIPS.
	It has C oriented instructions and supports compiling from GNU compilers.
	Apart from standard 16 bits core unit Digital Signal Processor brings very high performance for computation,
	for example FFT or other signal processing.
	Programs can be downloaded to flash memory with 144 kB and
	 are loaded to RAM memory according to the selector position.
	E-Puck's RAM memory has only 8 kB. A currently running program and all its data are placed in RAM memory.
	
	Communication with other devices is provided by Bluetooth and RS232 serial interface.
	Both, Bluetooth and RS232, can be used to download programs to e-Puck's flash memory.
	In addition, Bluetooth can be used to communicate with other e-Pucks or with a computer
	using {\it BTCom}. 
	The counter part of {\it BTCom} on computer is connected to the serial port.


