%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                    %%%
%%% ©ablona bakaláøské práce na MFF UK %%%
%%%                                    %%%
%%% (c) Franti¹ek ©trupl, 2005         %%%
%%%                                    %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% POZOR: Úprava bakaláøské práce je závislá rovnì¾ na volbì jednostranného resp. oboustranného tisku.
%%%        Bli¾¹i informace naleznete v dokumentu Úprava bakaláøské práce, který se nalézá na adrese:
%%%        http://www.mff.cuni.cz/studium/obecne/bplayout/pok12mo4.pdf

\documentclass[12pt,notitlepage]{report}
%\pagestyle{headings}
\pagestyle{plain}

\frenchspacing % aktivuje pou¾ití nìkterých èeských typografických pravidel

\usepackage[latin2]{inputenc} % nastavuje pou¾ité kódování, u¾ivatelé Windows zamìní latin2 za cp1250
%\usepackage{czech}
\usepackage[czech]{babel}
\usepackage{a4wide} % nastavuje standardní evropský formát stránek A4
%\usepackage{index} % nutno pou¾ít v pøípadì tvorby rejstøíku balíèkem makeindex
%\usepackage{fancybox} % umo¾òuje pokroèilé rámeèkování :-)
\usepackage{graphicx} % nezbytné pro standardní vkládání obrázkù do dokumentu

\usepackage[left=3cm]{geometry} % nastavení dané velikosti okrajù

%\newindex{default}{idx}{ind}{Rejstøík} % zavádí rejstøík v pøípadì pou¾ití balíku index

\title{Object Oriented Library for Controlling an e-Puck Robot}   % tyto dvì polo¾ky jsou zde v podstatì formálnì, ve skuteènosti nejsou nikde 
\author{Ondøej Plátek} % dále v dokumentu pou¾ity

%\date{}

\begin{document}
\chapter{Available software}
	\section{Introduction}
	In this chapter we shortly introduce software, which helps programmer
	to debug and develop a program, which control robot. Only simulators will be introduced, because 
	there is no available program or library for e-Puck like Elib,
	which controls e-Puck remotely and asynchronously from PC.
	All introduced simulators run at PC and allow user to control and program virtual copy of robot.
	Most of the simulators simplify physics and construction, although they try to simulate sensors
	and actuators as faithfully as possible. Some of the simulators support porting the program,
	which has been created for the virtual robot, to real robot.
	\\
	Both, Elib and the simulators, benefit from moving the debugging environment from the robot to PC.
	PC has enough resources at disposal and the code can be debugged much more comfortably. Debugging
	tools depends on the structure of simulator and a programming language, which is used by the simulator.
	Both attitudes allows to stop the program at any time.
	\\
	Elib uses $C\#$, which can be develop in Microsoft Visual Studio or MonoDevelop Studio.
	Both studios support many features like setting breakpoints with conditions, call stack table,
	manually switching between threads and many more.
	\\
	The simulators have several advantages over robot remote control.
	Simulator provides for developer a consistent behaviour of virtual robot.
	Programmer has for example under control battery level. The robot also measures in same situations
	identical values from its sensors. 
	\\
	Real robot is strongly influenced by the state of battery
	and measures a completely different values in the same situation.
	 Great advantage of simulator is that it can not only stop 
	 the program, but also freeze the simulated environment. 
	\\
	Simulators also do not need a real robot for testing a program, which is extremely
	useful in class with many students, where is only a few robots available.
	Some of the simulators are completely free, which is certainly nice quality.
	\\
	On the other hand a remote control of a robot has some benefits too. 
	Lot of unpredictable situations are not simulated in simulators, although 
	many simulators have the option to involve a randomness. The robot usually
	needs to be tested in real environment too.
	\\
	Because simulators use a physical and graphical engines, they are demanding applications
	for PC. Remote control can be simple fast application.
	\\
	Last but not least advantage of a remote control of robot is fact,
	 that playing with real robot is more fun. 	
	It motivates the students to solve the real problems including hardware set. Playing with robot brings
	 questions how to cope with a changing environment or what to do with an inaccurate odometry.
	\\
	In these chapter following simulators will be discussed:
	\begin{itemize}
	\item Microsoft Robotics Developer Studio is complex tool based on message sending.
	\item Player/Stage (Gazebo) is an open source project developed in C++, which is widely used.
	\item Pyrobot(Pyro) is also an open source project written in Python. It supports abstract
	interfaces for devices convenient for e-Puck.
	\item Enki is a fast simulation tool for big population of robots.
	\item Commercial simulator Webots is the only simulator, which almost fully support e-Puck.
	\end{itemize}
	%todo citation needed in all lists
\section{Microsoft Robotics Developer Studio (MSRS)}
	MSRS is set of tools including service-oriented runtime, simulator, Visual Programming Language (VPL),
	tutorial and examples. MSRS is based on .Net and its runtime introduces following new technologies:
	\begin{itemize}
	\item Concurrency and Coordination Runtime(CCR), which makes asynchronous programming easier.
	\item Decentralized Software Services(DSS) monitor services in real time for developers.
	\item Common Language Runtime (CLR) 2.0 is underlying runtime, 
		which interprets compiled code from any .Net language.
	\end{itemize}	
	MSRS offers VPL, which is convenient for beginners or for building a structure of a new application.
	VPL can be later easily transformed to any .Net language.
	\\
	Visual Simulation Environment is based on AGEIA PhysX engine. It enables a 3D simulation.
	\\
	In general MSRS is rich set of tools, which introduce many nice features including
	technologies for easy asynchronous and concurrent programming. Especially CCR really helps
	a programmer with concurrent programming. 
	\\
	There are enough tutorial for beginners written
	in $C\#$, Visual Basic or Python. (Iron Python is implementation of Python for .Net.)
	On the other hand so many new features and a lot of tools is a big overhead for students.
	MSRS tutorials and documentation has been rapidly improved in last year, but still some parts
	of the documentation are not easy accessible.
	\\
	MSRS runs only under Microsoft Windows operation systems and for a non commercial development
	it is for free. Neither e-Puck nor any similar robot is supported in latest version of MSRS.
	\\
	The studio would be good solution for e-Puck, because it meets the requirements for Elib.
	On the other hand it is a huge, not portable, complicated environment.
	Due to its complexity it discourages students from learning it.
\section{Player/Stage (Gazebo)}
	Player is a network server running on robot and sending sensors values from robot and
	accepting commands from clients. Player has a codified interface. Player translates
	commands from interface to implementation of actuators and translate values back from robot sensors 
	to Player interface. Unfortunately Player is not implemented for e-Puck
	and for its devices.
	\\
	Stage is a 2D simulator, which is reflecting the sensors and actuators using interface defined by Player.
	It simulates a population of mobile robots and therefore it is possible to use Player 
	in multi agent systems.
	The sensors are very simplified, because all sensors of the same type 
	have to implement the same interface. For example a sonar from Khephera and sonar from Aibo robot
	has the same interface and Staqe treats them equally. 
	Such simplification allows to Stage be much quicker.
	\\
	Gasebo is 3D simulator. Like Stage Gazebo simulates a big population of robots.
	It uses a rigid-body physics like Stage too.
	\\
	A whole project is open source and it is designed for POSIX-compatible systems.
	C/C++, Python, Java are main supported languages.
	Player/Stage is one of the most frequently used simulators in research and is good for education purposes.
	A Khephera robot, which is similar to e-Puck is supported.
	It indicates, that it is worth to think about Player/Stage, if simulator for e-Puck is wanted.
\section{Pyrobot(Pyro)}
	Pyro is a shortcut from Python Robotics. It is an open source project written in Python. Pyrobot is simulator
	from Pyro project and the name Pyrobot replaced the name of the project Pyro. 
	Pyro abstracts interfaces of robot and
	its devices, so each robot supported by Pyro can be treated equally.
	Pyro let you choose from list of simulators including Player/Stage, Gazebo, Pyrobot, Khephera simulator
	and Robocup Soccer.
	After selecting simulator, there are available different worlds according selected simulator.
	A simulated robot can be driven from command line or be controlled via loaded behaviour.
	\\
	The simulated robot can be easily replaced with real robot, if programs for controlling the real 
	and virtual robot are implemented. 	
	Controlling the real robot is done via communication with program running on the robot similar 
	to BTCom program on e-Puck.
	\\
	Khephera robot is supported. There is an implementation for e-Puck, which can simulate the robot
	and connect to real robot, but the implementation does not support asynchronous operations and does 
	not implement all sensors and actuators.
	\\
	For using Pyro as simulator for Elib it is needed to be created a bridge, 
	similar to which has already been implemented in e-Puck implementation.
	See future works (todo) for more information.	
	\\
	Pyro project was abandoned in year 2007 and no further development is planned.
	Advantage of project is its portability due to usage of Python and open source licence.
\section{Enki & Aseba}
	Enki is library, which can simulate several robots hundred times faster than real time.
	It is written in C++ and still is in development. It supports e-Puck including e-Puck's camera. 
	\\
	The e-Puck can be control via Aseba. Aseba is a tool for robot control.
	It introduce new scripting language. 
	The robots can be control only by the scripting language, which is the main
	drawback of Enki, because the scripting language is very simple.
	In addition Enki does not support remote control of robots and also a compiling program
	from the scripting language to e-Puck's microchip is impossible.
\section{Webots}
	Webots is simulation and prototyping tool for mobile robots. Only the simulation part will be described,
	because virtual e-Puck robot already has been designed.
	\\
	Webots divides the development in four stages. The first, designing a virtual robot, can be skipped,
	because the developers of Webots prepared for us finished virtual e-Puck.
	Second phase consist of programming robot, which includes getting values from sensors, 
	deciding what to do and	sending commands to robot. 
	Third phase is the simulation and fourth stage consist of loading a compiled program to e-Puck.
	\\
	Webots also have the option of controlling e-Puck via Bluetooth.
	Robot can be programmed with graphical programming language for beginners or with C/C++ or Java 
	but only C/C++ can be compiled for real e-Puck.
	\\
	The simulated worlds are created and saved in VRML and can be ported from 3D Max, Cinema 4D.
	\\
	Webots is complex environment, in which is easy to program, compile and load program to e-Puck.
	The simulation runs in 3D and Webots supports all sensors and actuators except microphones.
	Big advantage is that the real robot can be controlled by Bluetooth or programmed by loading a compiled program.
	Webots is also well documented and good tutorial are available.
	The big problem of Webots is its price and its annual fees.
	\\

 %todo citation
	
\end{document}

